Data Anonymization for Open Science: A Case Study

### 2025-05-14 submission_materials/article.tex
Update article.tex
added script to compare datasets, a dataset description and slightly extended the manuscript to comply with some of the revision requests.

\hl{The quality and utility of anonymized data has long been studied. It is common to distinguish between general-purpose and special-purpose (or workload-aware) approaches for quantifying anonymized data quality. General-purpose approaches include information loss metrics, like granularity reduction} \mycite{10} or record fidelity \mycite{11}, \hl{as well as resemblance metrics, which can be univariate, e.g. comparing distributions of variables in unprotected vs. anonymized data} \mycite{12}, \hl{or multivariate, e.g. comparing differences in correlation matrices} \mycite{13}. \hl{Another approach is to study how well classifiers can distinguish unprotected and anonymized records} \mycite{14}. \hl{All three anonymization methods in this paper, along with 13 other methods, were compared across a risk assessment as well as seven quality metrics including univariate, 2-marginal, and 3-marginal, linear regression, propensity mean squared error, principal component analysis, and data inconsistencies}\mycite{15}.

\hl{Other prior work proposes that the anonymization mechanisms themselves should be workload aware: they take into account what analysis is going to be done after anonymization} \mycite{16}\mycite{17}. \hl{The quality evaluations in this work are therefore naturally based on the workload analysis.}

\hl{All of the quality measures in this prior work, however, are generic in that they do not take into account a specific scientific goal. These generic measures are useful for comparing the quality of different anonymization techniques generally, but they fall short of predicting whether any given scientific conclusion will be correct or not. By contrast, our paper measures quality based on specific scientific conclusions.}

@@

@@

The original and anonymized data for base-tables 1 and 2 explored the types of commuting in school children and the commute distances traveled. These are given in this paper's Tables 1 and 2 respectively\mycite{18}.

Among the three anonymization methods, SynDiffix most closely reproduced original counts and percentages for cross-tabulation of commuting modes (Table 1), with only one deviating record in the cross table. It is closely followed by ARX which showed up to 18 deviations, while SDV provides large differences of up to 179 different commute mode counts in the cross table. Nevertheless, the basic description of these results in original paper, i.e. that active commuting is more frequent in the direction from school to home than in other direction, still holds, but this may be a coincidence in case of SDV as it substantially over- or underestimates the frequencies for both modes of active commuting (walking and wheels). Note also that the anonymized table of ARX contains 1.4\% fewer rows (703 versus 713) due to suppression of rare records.

Similar results were found for commuting distance (Table~2). Here also SDV performs badly, hugely over- or underestimating the distance in most of the table cells, both in central tendency (median) and variability (IQR) of the data. SynDiffix and ARX perform much better, but still with large errors in some of the table cells, especially in cells with low frequencies, e.g. SynDiffix substantially underestimates the median distance of car commuting from school to home (3615$\rightarrow$2584) and ARX overestimates the distance of wheels commuting in same direction (1444$\rightarrow$2235). The verbal description of this results from original paper, i.e. that active commuting groups (Walk, Wheels) typically live
close to school, still holds in case of SynDiffix and ARX, but not in case of SDV.

%\begin{quotation}
%\color{blue}
%
%Participants more often used active commuting from
%school (56\%) than to school (44\%). Males and females
%were choosing active commuting equally often, though
%the notable difference was that females preferred to walk and males preferred to use wheels transport. The Walk
%group had the lowest and Public group had the highest
%median distance from home to school (Table~2).
%
%Note that, commuting to school may represent a relatively small amount of total PA of Slovenian children
%compared to their all other daily physical activities, as
%the active commuting groups (Walk Wheels) typically
%live close to school (Table 2).
%
%\end{quotation}

@@

      \bottomrule
      \end{tabular}
      \end{small}
      \caption{Base-table 2 from the original paper showing the counts and distances in meters (median and IQR) for the original data and the three anonymization methods. Each group of four presents the data in order of Original, ARX, SDV, and SynDiffix. The bold/italic fonts for counts (N) are as described for Table 1. Distance and IRQ are \textbf{bold font} when the relative error is greater than 30\%, and \textit{italics} when the relative error is greater than 15\%.  Note that the original distances median and IQR don't perfectly match those of the original Table 2 because of differences in the way median and IQR were calculated (Python versus R).}
      \label{tab:table2}
      \end{center}
      \end{table}
      \setlength{\fboxsep}{3pt}

@@

The absolute errors of the cross-table counts from Table 1 and the distance values in Tables from 2 are visualized in the boxplots of Figure 1.

@@

      \bottomrule
      {\footnotesize * p $\leq$ 0.05, \quad** p $\leq$ 0.01, \quad*** p $\leq$ 0.001}
      \end{tabular}
      \end{small}
      \caption{Part 1 (of 2) of the original paper's Table 3 showing the parameters (regression coefficients) of the linear model for prediction of VO2max by group and distance. \textbf{Bold font} indcates that the anonymized entry is non-significant where the original data is significant or vice versa. Each group of four presents the data in order of Original, ARX, SDV, and SynDiffix. 
      }
      \label{tab:table3a}
      \end{center}
      \end{table}
      \setlength{\fboxsep}{3pt}

@@

      \bottomrule
      {\footnotesize * p $\leq$ 0.05, \quad** p $\leq$ 0.01, \quad*** p $\leq$ 0.001}
      \end{tabular}
      \end{small}
      \caption{Part 2 (of 2) of the original paper's Table 3 showing the parameters (regression coefficients) of the linear model for prediction of VO2max by group and distance. \textbf{Bold font} indcates that the anonymized entry is non-significant where the original data is significant or vice versa. Each group of four presents the data in order of Original, ARX, SDV, and SynDiffix. 
      }
      \label{tab:table3b}
      \end{center}
      \end{table}
      \setlength{\fboxsep}{3pt}

@@

The original and anonymized data for base-table 3 are given is this paper's Tables~3 and 4 (the data is spread over two tables for formatting purposes)\mycite{18}. The primary purpose of base-table 3 is to indicate statistical significance. As such, those entries where the original data is significant and the anonymized data is not, or vice versa, are highlighted in red. In total, there are 3 such mismatches for ARX, 8 for SDV, and 5 for SynDiffix.

Several differences were found in statistics of linear prediction models based on original data and the data produced by all three anonymization methods (Tables~3 and 4). In the original models predicting VO2max, the constant (intercept), the predictors Gender and MVPA, and the derived Car × Gender interaction term were statistically significant in both directions of commuting. This also holds for all parameters in the SynDiffix and ARX models, except for the Car x Gender parameter in the SynDiffix model, which is not significant in one of the directions of commuting. The Gender parameter in the SDV model was found to be non-significant, despite being highly significant (p$\leq$.001) in the three other models.

Note that only the SDV method provides estimate for Wheels x Males interaction parameter in the school to home direction models. ARX and SynDiffix suppressed this data as part of anonymization because there were so few datapoints.
%
The normalized error for coefficients in Tables 3 and 4 is illustrated in Figure 2.

@@

% In the following table, use:
%    \cmark for the green checkmark
%    \omark for the yellow circle
%    \xmark for the red X
\begin{table}
\begin{center}
\begin{small}
\begin{tabularx}{\textwidth}{Xlll}
\toprule
\textbf{Main policy conclusion}  & ARX & SDV & SynDiffix \\
Children driven by car who live
within wheels or walk distance from school should be targeted by interventions promoting active transport & \cmark & \xmark & \cmark \\
\midrule
\textbf{Descriptive analytics results}  & ARX & SDV & SynDiffix \\
Overall, 43\% of the participants reported active commuting modes to and from school and an additional 13\%
only in one direction & \cmark & \xmark & \cmark \\
Participants more often used active commuting from school (56\%) than to school (44\%) & \cmark & \cmark & \cmark \\
Males and females were choosing active commuting equally often & \cmark & \cmark & \cmark \\
Females preferred to walk and males preferred to use wheels transport & \cmark & \xmark & \cmark \\
The Walk group had the lowest and Public group had the highest median distance from home to school & \cmark & \xmark & \cmark \\
The active commuting groups (Walk Wheels) typically live close to school & \cmark & \xmark & \cmark \\
\midrule
\textbf{CRF-related analytic results}  & ARX & SDV & SynDiffix \\
When commuting from school to home, both the main effect of commuting group and its interaction with
distance were significant & \cmark & \xmark & \cmark \\
The Car group had a significantly lower predicted value of CRF & \cmark & \xmark & \cmark \\
The Car group was the only one where distance of travel was related to CRF having significant difference to the reference (Walk) group; the participants with larger car travel distance had higher predicted VO2max & \cmark & \xmark & \omark \\
Overall interaction of commuting group with gender was not significant  & \omark & \cmark & \omark \\
The largest sample difference between males and females was observed in the reference group (Walk), but it was only significant in the Car group & \omark & \omark & \omark \\
When commuting from home to school neither main effect of commuting group, nor it’s interactions with
gender and commuting distance were significant & \omark & \cmark & \cmark \\
\midrule
\textbf{Main scientific results}  & ARX & SDV & SynDiffix \\
The main effect of commuting group on CRF and its interaction with distance were significant in the
direction from school to home, but not in the opposite direction & \omark & \xmark & \omark \\
Predicted differences in CRF between commuting groups were moderate and generally higher in males than in females & \cmark & \xmark & \cmark \\
When comparing commuting group median distance from school to home, males driven by car had around 4 ml/
min/kg lower predicted CRF than those who walked or used wheels commuting (e.g., bicycle, skateboard) & \omark & \xmark & \omark \\

@@

Table 5 summarizes the ability of each anonymization method to produce the same analytic conclusions as those of the base paper. It presents the set of statements made in the base paper, and evaluates whether the statement would be supported (\cmark), negated (\xmark), or neither supported nor negated (\omark).

@@

\hl{For our measure, we removed 100 records from the dataset to establish the baseline, and re-anonymized the remaining 613 records. We assumed that the attacker knows the victim's gender, age, and commuting distance and mode in both directions. These are all potentially publicly-known attributes. The unknown attributes being predicted are V02max and MVPAsqrt. In evaluating SynDiffix, the two tables with only the known attributes are unknown attribute were used. Table} 6 \hl{presents the Risk scores and the 95\% confidence intervals. In all cases, even the high confidence bound is well below the strong privacy Risk threshold of 0.5.}

\begin{table}
\begin{center}
\begin{small}
\begin{tabular}{rcccc}
\toprule
Method & \multicolumn{2}{c}{VO2max} & \multicolumn{2}{c}{MVPAsqrt} \\ 
 & Risk & (CI low, CI high) & Risk & (CI low, CI high) \\ 
\midrule
ARX & 0.139 & (0.0, 0.338) & 0.131 & (0.023, 0.239) \\ 
SDV & 0.0 & (0.0, 0.123) & 0.031 & (0.0, 0.106) \\ 
SynDiffix & 0.017 & (0.0, 0.248) & 0.042 & (0.0, 0.123) \\ 
\bottomrule
\end{tabular}
\end{small}
\caption{Privacy risk scores and 95\% confidence intervals. Any risk score below 0.5 can be regarded as having very strong anonymity.}
\label{tab:risk_eval}
\end{center}
\end{table}

@@

Comparison of the VO2max data plots corresponding to the base-figure from the original study.

@@

(c) SynDiffix Plot

>>>>

\hl{The quality and utility of anonymized data has long been studied. It is common to distinguish between general-purpose and special-purpose (or workload-aware) approaches for quantifying anonymized data quality. General-purpose approaches include information loss metrics, like granularity reduction} \mycite{10} or record fidelity \mycite{11}, \hl{as well as resemblance metrics, which can be univariate, e.g. comparing distributions of variables in unprotected vs. anonymized data} \mycite{12}, \hl{or multivariate, e.g. comparing differences in correlation matrices} \mycite{13}. \hl{Another approach is to study how well classifiers can distinguish unprotected and anonymized records} \mycite{14}. The three presented anonymization methods have been previously explored regarding the general-purpose utility of the protected data \mycite{15}, whereas workload-aware utility analysis was not yet explored for all methods. 
\hl{Other prior work proposes that even the anonymization mechanisms themselves should be workload aware: they take into account what analysis is going to be done after anonymization} \mycite{16}\mycite{17}. \hl{The quality evaluations in this work are therefore naturally based on the workload analysis, which is a novelty of this paper.}

\hl{All of the quality measures in this prior work, however, are generic in that they do not take into account a specific scientific goal. These generic measures are useful for comparing the quality of different anonymization techniques generally, but they fall short of predicting whether any given scientific conclusion will be correct or not. By contrast, our paper measures quality based on specific scientific conclusions, which is another novelty of this study. Here we directly compare the ability of the anonymization methods to preserve the data characteristics that lead to the scientific conclusions.}

\hl{Although the data quality and utility was previously thoroughly explored, previous research on data privacy risks associated with non-traditional anonymization methods and synthetic data generation is rather sparse [BAYREM]. Few frameworks exist that can estimate residual privacy risks on anonymized and synthetic data alike. One of such framework is the Anonymeter framework by Giomi et al \mycite{20}, originally proposed for large, synthetic datasets to evaluate risk of linkage, attribute inference and record singling-out. 
}

@@

\subsection*{Dataset comparison}

\setlength{\fboxsep}{0pt}
\begin{table}
      \begin{center}
      \begin{small}
\begin{tabular}{lllll}
\toprule
 & \textbf{Original}  & \textbf{ARX} & \textbf{SDV} & \textbf{SynDiffix} \\
 & n=713 & n=703  & n=713  & n=715   \\
\midrule
Age & 13.34 (0.91) & 13.35 (0.81)  & 13.48 (1.02)  & 13.33 (0.90) \\
MVPA & 30.03 (9.35) & 30.04 (7.63) & 34.06 (11.65)* & 30.07 (9.18) \\
VO$_2$max & 48.22 (6.3) & 48.29 (5.52)  & 45.53 (7.29)*  & 48.23 (6.34) \\
Distance to school & 3319 (3839) & 3304 (3467) & 7499 (6599)* & 3286 (3641) \\
Distance from school & 3287 (3832) & 3275 (3463) & 5233 (6664)* & 3252 (3609) \\
\bottomrule
\end{tabular}
      \end{small}
      \caption{Descriptive statistics of continuous variables expressed as mean and standard deviation. * statistically significant differences based on two-sided independent t-tests and alpha-level of 0.005.}
      \label{tab:table1}
      \end{center}
      \end{table}
	\setlength{\fboxsep}{3pt}

Descriptive univariate statistics as mean and standard deviations are provided in Table 1, showing overall similar distributions and no statistical significant difference to the original data. The data generated by SDV showed differences in four out of five variables with p-values below .001. The p-value for the Age variable was 0.006, which is very close to the chosen alpha-level of 0.005. Looking at the difference in mean values, we see huge deviations in the SDV dataset for both commute distances, potentially indicating general flaws in the synthetic data generation used here.

@@

The original and anonymized data for base-tables 1 and 2 explored the types of commuting in school children and the commute distances traveled. These are given in this paper's Tables 2 and 3 respectively\mycite{18}.

Among the three anonymization methods, SynDiffix most closely reproduced original counts and percentages for cross-tabulation of commuting modes (Table 2), with only one deviating record in the cross table. It is closely followed by ARX which showed up to 18 deviations, while SDV provides large differences of up to 179 different commute mode counts in the cross table. Nevertheless, the basic description of these results in original paper, i.e. that active commuting is more frequent in the direction from school to home than in other direction, still holds, but this may be a coincidence in case of SDV as it substantially over- or underestimates the frequencies for both modes of active commuting (walking and wheels). Note also that the anonymized table of ARX contains 1.4\% fewer rows (703 versus 713) due to suppression of rare records.

Similar results were found for commuting distance (Table~3). Here also SDV performs badly, hugely over- or underestimating the distance in most of the table cells, both in central tendency (median) and variability (IQR) of the data. SynDiffix and ARX perform much better, but still with large errors in some of the table cells, especially in cells with low frequencies, e.g. SynDiffix substantially underestimates the median distance of car commuting from school to home (3615$\rightarrow$2584) and ARX overestimates the distance of wheels commuting in same direction (1444$\rightarrow$2235). The verbal description of this results from original paper, i.e. that active commuting groups (Walk, Wheels) typically live
close to school, still holds in case of SynDiffix and ARX, but not in case of SDV.

%\begin{quotation}
%\color{blue}
%
%Participants more often used active commuting from
%school (56\%) than to school (44\%). Males and females
%were choosing active commuting equally often, though
%the notable difference was that females preferred to walk and males preferred to use wheels transport. The Walk
%group had the lowest and Public group had the highest
%median distance from home to school (Table~3).
%
%Note that, commuting to school may represent a relatively small amount of total PA of Slovenian children
%compared to their all other daily physical activities, as
%the active commuting groups (Walk Wheels) typically
%live close to school (Table 2).
%
%\end{quotation}

@@

      \bottomrule
      \end{tabular}
      \end{small}
      \caption{Base-table 2 from the original paper showing the counts and distances in meters (median and IQR) for the original data and the three anonymization methods. Each group of four presents the data in order of Original, ARX, SDV, and SynDiffix. The bold/italic fonts for counts (N) are as described for Table 2. Distance and IRQ are \textbf{bold font} when the relative error is greater than 30\%, and \textit{italics} when the relative error is greater than 15\%.  Note that the original distances median and IQR don't perfectly match those of the base-table 2 because of differences in the way median and IQR were calculated (Python versus R).}
      \label{tab:table2}
      \end{center}
      \end{table}
      \setlength{\fboxsep}{3pt}

@@

The absolute errors of the cross-table counts from Table 2 and the distance values in Tables from 3 are visualized in the boxplots of Figure 1.

@@

      \bottomrule
      {\footnotesize * p $\leq$ 0.05, \quad** p $\leq$ 0.01, \quad*** p $\leq$ 0.001}
      \end{tabular}
      \end{small}
      \caption{Part 1 (of 2) of the base-table 3 showing the parameters (regression coefficients) of the linear model for prediction of VO$_2$max by group and distance. \textbf{Bold font} indcates that the anonymized entry is non-significant where the original data is significant or vice versa. Each group of four presents the data in order of Original, ARX, SDV, and SynDiffix. 
      }
      \label{tab:table3a}
      \end{center}
      \end{table}
      \setlength{\fboxsep}{3pt}

@@

      \bottomrule
      {\footnotesize * p $\leq$ 0.05, \quad** p $\leq$ 0.01, \quad*** p $\leq$ 0.001}
      \end{tabular}
      \end{small}
      \caption{Part 2 (of 2) of the original paper's Table 3 showing the parameters (regression coefficients) of the linear model for prediction of VO$_2$max by group and distance. \textbf{Bold font} indcates that the anonymized entry is non-significant where the original data is significant or vice versa. Each group of four presents the data in order of Original, ARX, SDV, and SynDiffix. 
      }
      \label{tab:table3b}
      \end{center}
      \end{table}
      \setlength{\fboxsep}{3pt}

@@

The original and anonymized data for base-table 3 are given is this paper's Tables~4 and 5 (the data is spread over two tables for formatting purposes)\mycite{18}. The primary purpose of base-table 3 is to indicate statistical significance. As such, those entries where the original data is significant and the anonymized data is not, or vice versa, are highlighted in red. In total, there are 3 such mismatches for ARX, 8 for SDV, and 5 for SynDiffix.

Several differences were found in statistics of linear prediction models based on original data and the data produced by all three anonymization methods (Tables~4 and 5). In the original models predicting VO$_2$max, the constant (intercept), the predictors Gender and MVPA, and the derived Car × Gender interaction term were statistically significant in both directions of commuting. This also holds for all parameters in the SynDiffix and ARX models, except for the Car x Gender parameter in the SynDiffix model, which is not significant in one of the directions of commuting. The Gender parameter in the SDV model was found to be non-significant, despite being highly significant (p$\leq$.001) in the three other models.

Note that only the SDV method provides estimate for Wheels x Males interaction parameter in the school to home direction models. ARX and SynDiffix suppressed this data as part of anonymization because there were so few datapoints.
%
The normalized error for coefficients in Tables 4 and 5 is illustrated in Figure 2.

@@

% In the following table, use:
%    \cmark for the green checkmark
%    \omark for the yellow circle
%    \xmark for the red X
\begin{table}
\begin{center}
\begin{small}
\begin{tabularx}{\textwidth}{Xlll}
\toprule
\textbf{Main policy conclusion}  & ARX & SDV & SynDiffix \\
Children driven by car who live
within wheels or walk distance from school should be targeted by interventions promoting active transport & \cmark & \xmark & \cmark \\
\midrule
\textbf{Descriptive analytics results}  & ARX & SDV & SynDiffix \\
Overall, 43\% of the participants reported active commuting modes to and from school and an additional 13\%
only in one direction & \cmark & \xmark & \cmark \\
Participants more often used active commuting from school (56\%) than to school (44\%) & \cmark & \cmark & \cmark \\
Males and females were choosing active commuting equally often & \cmark & \cmark & \cmark \\
Females preferred to walk and males preferred to use wheels transport & \cmark & \xmark & \cmark \\
The Walk group had the lowest and Public group had the highest median distance from home to school & \cmark & \xmark & \cmark \\
The active commuting groups (Walk Wheels) typically live close to school & \cmark & \xmark & \cmark \\
\midrule
\textbf{CRF-related analytic results}  & ARX & SDV & SynDiffix \\
When commuting from school to home, both the main effect of commuting group and its interaction with
distance were significant & \cmark & \xmark & \cmark \\
The Car group had a significantly lower predicted value of CRF & \cmark & \xmark & \cmark \\
The Car group was the only one where distance of travel was related to CRF having significant difference to the reference (Walk) group; the participants with larger car travel distance had higher predicted VO$_2$max & \cmark & \xmark & \omark \\
Overall interaction of commuting group with gender was not significant  & \omark & \cmark & \omark \\
The largest sample difference between males and females was observed in the reference group (Walk), but it was only significant in the Car group & \omark & \omark & \omark \\
When commuting from home to school neither main effect of commuting group, nor it’s interactions with
gender and commuting distance were significant & \omark & \cmark & \cmark \\
\midrule
\textbf{Main scientific results}  & ARX & SDV & SynDiffix \\
The main effect of commuting group on CRF and its interaction with distance were significant in the
direction from school to home, but not in the opposite direction & \omark & \xmark & \omark \\
Predicted differences in CRF between commuting groups were moderate and generally higher in males than in females & \cmark & \xmark & \cmark \\
When comparing commuting group median distance from school to home, males driven by car had around 4 ml/
min/kg lower predicted CRF than those who walked or used wheels commuting (e.g., bicycle, skateboard) & \omark & \xmark & \omark \\

@@

Table 6 summarizes the ability of each anonymization method to produce the same analytic conclusions as those of the base paper. It presents the set of statements made in the base paper, and evaluates whether the statement would be supported (\cmark), negated (\xmark), or neither supported nor negated (\omark).

@@

\hl{For our measure, we removed 100 records from the dataset to establish the baseline, and re-anonymized the remaining 613 records. We assumed that the attacker knows the victim's gender, age, and commuting distance and mode in both directions. These are all potentially publicly-known attributes. The unknown attributes being predicted are VO$_2$max and MVPA. In evaluating SynDiffix, the two tables with only the known attributes are unknown attribute were used. Table} 7 \hl{presents the Risk scores and the 95\% confidence intervals. In all cases, even the high confidence bound is well below the strong privacy Risk threshold of 0.5.}

\begin{table}
\begin{center}
\begin{small}
\begin{tabular}{rcccc}
\toprule
Method & \multicolumn{2}{c}{VO$_2$max} & \multicolumn{2}{c}{MVPA} \\ 
 & Risk & (CI low, CI high) & Risk & (CI low, CI high) \\ 
\midrule
ARX & 0.139 & (0.0, 0.338) & 0.131 & (0.023, 0.239) \\ 
SDV & 0.000 & (0.0, 0.123) & 0.031 & (0.0, 0.106) \\ 
SynDiffix & 0.017 & (0.0, 0.248) & 0.042 & (0.0, 0.123) \\ 
\bottomrule
\end{tabular}
\end{small}
\caption{Privacy risk scores and 95\% confidence intervals. Any risk score below 0.5 can be regarded as having very strong anonymity.}
\label{tab:risk_eval}
\end{center}
\end{table}

@@

Comparison of the VO$_2$max data plots corresponding to the base-figure from the original study.

@@

(d) SynDiffix Plot
